---
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# eurolig

[![lifecycle](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)

The eurolig package provides a set of tools to obtain and analyze basketball data from the Euroleague.

## Overview

* API wrapper around the official website of the [Euroleague](https://www.euroleague.net/).

* Functions to obtain stats, play-by-play data and shot location data in a tidy format within R.

* Tools to analyze and visualize the data.


## Installation

```{r, eval = FALSE}
# install.packages("devtools")
devtools::install_github("solmos/eurolig")
```

## Usage

Let's say we want to analyze play-by-play data from the Championship game of the 2017-2018 season between Real Madrid and Fenerbache Dogus Istanbul. We can obtain the data by entering the game code and the season to the function `extract_pbp()`. The game code for the game can be found in the game's [URL](https://www.euroleague.net/main/results/showgame?gamecode=260&seasoncode=E2017).

```{r}
library(eurolig)
pbp <- extract_pbp(game_code = 260, season = 2017)
str(pbp)
```

