---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
)
```
# eurolig

[![lifecycle](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)

The eurolig package provides a set of tools to obtain and analyze basketball generated data from the Euroleague.

## Overview

* API wrapper around the official website of the [Euroleague](https://www.euroleague.net/).

* Functions to obtain stats, play-by-play data and shot location data in a tidy format within R.

* Tools to analyze and visualize the data.


## Installation

```{r, eval = FALSE}
# install.packages("devtools")
devtools::install_github("solmos/eurolig")
```


## Example

Let's say we want to analyze play-by-play data from the Championship game of the 2017-2018 season between Real Madrid and Fenerbache Dogus Istanbul. We can obtain the data by entering the game code and the season to the function `extractPbp()`. The game code for the game can be found in the game's [URL](https://www.euroleague.net/main/results/showgame?gamecode=260&seasoncode=E2017).

```{r}
library(eurolig)
pbp <- extractPbp(game_code = 260, season = 2017)
pbp
```

From these data we can extract information about the assists in that game from, say, Real Madrid (MAD):
```{r}
assists <- getAssists(pbp, team = "MAD")
assists
```

Shot location data can also be retrieved and visualized:
```{r}
shots <- extractShots(260, 2017)
plotShotchart(shots)
```

